---
title: "ANOVA testing Notebook"
output:
  prettydoc::html_pretty:
    toc       : true
    df_print  : paged
    theme     : Cayman
    fig_retina : true
---

In this notebook, we produce ANOVA analysis of several data sets for validation purposes. We evaluate separetely each type of ANOVA. As discissued previously ANOV type I is optional and can be ignore. The first part tests balanced data whereas the second part test unbalanced data. In the MIP we have to deal with unbalanced data.  
<br>
the results of anova must contain at least F-statistic, p_value, eta-squared and partial eta-squared for publications. But it is usual to display sum of the squares and the mean sum of the squares in a table format. 


# Local calculation (single node)
All this test are runing with a single dataset. 

## Setup environment and load dataset

The dependent variable is centered to calculate eta squared and partial eta squared. Centering has no influence on the calculation of other items, p, F, sumsq, ...

```{r message = F}
library(tidyverse)
library(jmv)

# load balanced data 
rawDataB <- read_delim("./Data/data_ANOVA_Balanced_with_inter_V1V2.csv", ";") %>%
  mutate(
    var_I1 = as.factor(var_I1),
    var_I2 = as.factor(var_I2),
    var_I3 = as.factor(var_I3)
    ) 

rawDataB
```

## Tests With Balanced data sets

In this series we use spacific balanced dataset and evaluate all the three ANOVA tyypes. 

- Test the 3 main effects and all the interactions with type I ANOVA. 

```{r}
jmv::ANOVA(
  formula   = var_D ~ var_I1*var_I2*var_I3, 
  data      = rawDataB,
  ss        = 1,
 effectSize = c('eta', 'partEta')
  )
```

- Test the 3 main effects and all the interactions with type II ANOVA. 

```{r}
# create linear model for balanced data
jmv::ANOVA(
  formula   = var_D ~ var_I1*var_I2*var_I3, 
  data      = rawDataB,
  ss        = 2,
 effectSize = c('eta', 'partEta')
  )
```

- Test the 3 main effects and all the interactions with type III ANOVA. 

```{r}
jmv::ANOVA(
  formula   = var_D ~ var_I1*var_I2*var_I3, 
  data      = rawDataB,
  ss        = 1,
  effectSize = c('eta', 'partEta')
  )
```

Remark : In a balanced data set, the result of Anova does not depend on how the sums of the squares are calculated. Hence, the three results are equal!
It is not the case in unbalanced data sets.



## With unbalanced data sets

In this notebook, we produce ANOVA analysis of several data sets for validation purposes. We evaluate separetely each type of ANOVA. As discissued previously ANOV type I is optional and can be ignore.

Import Balanced data
```{r message = F}
rawDataUB <- read_delim("./Data/data_ANOVA_Unbalanced_with_inter_V1V2.csv", ";") 
rawDataUB <- rawDataUB %>%
  mutate(
    var_I1 = as.factor(var_I1),
    var_I2 = as.factor(var_I2),
    var_I3 = as.factor(var_I3)
    )
rawDataUB
```

- Test the 3 main effects and all the interactions with type I ANOVA. 

```{r}
# create linear model for balanced data
jmv::ANOVA(
  formula   = var_D ~ var_I1*var_I2*var_I3, 
  data      = rawDataUB,
  ss        = 1,
  effectSize = c('eta', 'partEta')
  )
```

- Test the 3 main effects and all the interactions with type II ANOVA. 

```{r}
jmv::ANOVA(
  formula   = var_D ~ var_I1*var_I2*var_I3, 
  data      = rawDataUB,
  ss        = 2,
  effectSize = c('eta', 'partEta')
  )
```

- Test the 3 main effects and all the interactions with type III ANOVA. 

```{r}
jmv::ANOVA(
  formula   = var_D ~ var_I1*var_I2*var_I3, 
  data      = rawDataUB,
  ss        = 3,
  effectSize = c('eta', 'partEta')
  )
```




- Test the 3 main effects only with type III ANOVA. 

```{r}
# create linear model for balanced data with contrast == sum
jmv::ANOVA(
  formula   = var_D ~ var_I1 + var_I2 + var_I3, 
  data      = rawDataUB,
  ss        = 3,
  effectSize = c('eta', 'partEta')
  )
```



### With data similar to our target

We now test some 'real life' like configurations. We only test type II and III ANOVAs because in the MIP environment the data are very unbalanced and type I is inappropriate. 

```{r}
df0 <- read_csv("Data/dataset_0.csv") %>%
  mutate(
    agegroup                = as.ordered(agegroup), 
    alzheimerbroadcategory  = as.factor(alzheimerbroadcategory), 
    dataset                 = as.factor(dataset), 
    gender                  = as.factor(gender)
    )
df0
```



### ANOVA testing

- without interaction and 2 variables - type III

```{r}
jmv::ANOVA(
  formula   = lefthippocampus ~ alzheimerbroadcategory + gender, 
  data      = df0,
  ss        = 3,
  effectSize = c('eta', 'partEta')
  )
```

- without interaction and 2 variables - type II

```{r}
jmv::ANOVA(
  formula   = lefthippocampus ~ alzheimerbroadcategory + gender, 
  data      = df0,
  ss        = 2,
  effectSize = c('eta', 'partEta')
  )
```


- with interaction and 2 variables - type III

```{r}
jmv::ANOVA(
  formula = lefthippocampus ~ alzheimerbroadcategory * gender, 
  data = df0,
  ss = 3,
  effectSize = c('eta', 'partEta')
  )
```

- with interaction and 2 variables - type II

```{r}
jmv::ANOVA(
  formula = lefthippocampus ~ alzheimerbroadcategory * gender, 
  data = df0,
  ss = 2,
  effectSize = c('eta', 'partEta')
  )
```

- without interaction and 3 variables - type III

```{r}
jmv::ANOVA(
  formula = lefthippocampus ~ alzheimerbroadcategory + gender + agegroup, 
  data = df0,
  ss = 3,
  effectSize = c('eta', 'partEta')
  )
```

- without interaction and 3 variables - type II
  
```{r}
jmv::ANOVA(
  formula = lefthippocampus ~ alzheimerbroadcategory + gender + agegroup, 
  data = df0,
  ss = 2,
  effectSize = c('eta', 'partEta')
  )
```

- with 1 interaction and 3 variables - type III

```{r}
jmv::ANOVA(
  formula = lefthippocampus ~ alzheimerbroadcategory * gender + agegroup, 
  data = df0,
  ss = 3,
  effectSize = c('eta', 'partEta')
  )
```

- with 1 interaction and 3 variables - type II

```{r}
jmv::ANOVA(
  formula = lefthippocampus ~ alzheimerbroadcategory * gender + agegroup, 
  data = df0,
  ss = 2,
  effectSize = c('eta', 'partEta')
  )
```

- with full interaction and 3 variables - type III

```{r}
jmv::ANOVA(
  formula = lefthippocampus ~ alzheimerbroadcategory * gender * agegroup, 
  data = df0,
  ss = 3,
  effectSize = c('eta', 'partEta')
  )
```


- with full interaction and 3 variables - type II

```{r}
jmv::ANOVA(
  formula = lefthippocampus ~ alzheimerbroadcategory * gender * agegroup, 
  data = df0,
  ss = 2,
  effectSize = c('eta', 'partEta')
  )
```




