---
title: "PearsonCorrelation"
output:
  prettydoc::html_pretty:
    toc       : true
    df_print  : paged
    theme     : Cayman
    fig_retina : true
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup analysis environment

Load libraries "tidyverse" and "jmv" <br>


```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(jmv)
```

Load dataset "desd-synthdata.csv" and check number of rows = 1000.

```{r message=F}
rawData <- read_csv("Data/desd-synthdata.csv")
```

## Test the correlation between two highly correlated continuous variables

```{r echo=FALSE, message=FALSE}

df_PC1 <- rawData %>%
  select(
    righthippocampus,
    lefthippocampus) %>%
  drop_na() 
head(df_PC1)
```

```{r}
corrPears_1 <- jmv::corrMatrix(
    data = df_PC1,
    pearson = TRUE,
    vars = vars(lefthippocampus, righthippocampus),
    flag = TRUE,
    ci = TRUE,
    plots = TRUE,
    plotDens = TRUE,
    plotStats = TRUE)
corrPears_1$matrix
```

```{r}
corrPears_1$plot
```

## Test the correlation between two weakly correlated continuous variables


### Positive correlation

```{r echo=FALSE, message=FALSE}

df_PC2 <- rawData %>%
  select(
    lefthippocampus,
    opticchiasm
    ) %>%
  drop_na() 

head(df_PC2)
```

```{r}
corrPears_2 <- jmv::corrMatrix(
    data = df_PC2,
    pearson = TRUE,
    vars = vars(lefthippocampus, opticchiasm),
    flag = TRUE,
    ci = TRUE,
    plots = TRUE,
    plotDens = TRUE,
    plotStats = TRUE)
corrPears_2$matrix
```

```{r}
corrPears_2$plot
```


### Negative correlation
```{r echo=FALSE, message=FALSE}

df_PC2.1 <- rawData %>%
  select(
    subjectage,
    minimentalstate
    ) %>%
  drop_na() 

head(df_PC2.1)
```

```{r}
corrPears_2.1 <- jmv::corrMatrix(
    data = df_PC2.1,
    pearson = TRUE,
    vars = vars(subjectage, minimentalstate),
    flag = TRUE,
    ci = TRUE,
    plots = TRUE,
    plotDens = TRUE,
    plotStats = TRUE)
corrPears_2.1$matrix
```

```{r}
corrPears_2.1$plot
```

> Comment: We should not use Pearson correlation in this case since minimentalstate is not normaly distributed.

## Test the correlation between two weakly correlated continuous variables with different magnitude

### With data similar to MIP

```{r echo=FALSE, message=FALSE}

df_PC3 <- rawData %>%
  select(
    opticchiasm,
    subjectage
    ) %>%
  drop_na() 

head(df_PC3)
```

```{r}
corrPears_3 <- jmv::corrMatrix(
    data = df_PC3,
    pearson = TRUE,
    vars = vars(subjectage, opticchiasm),
    flag = TRUE,
    ci = TRUE,
    plots = TRUE,
    plotDens = TRUE,
    plotStats = TRUE)
corrPears_3$matrix
```

```{r}
corrPears_3$plot
```

### With extreme data 

```{r}
df_PC3.1 <- read_csv("Data/data_pr1.csv")
df_PC3.1
```
 
#### By changing the magnitude only
```{r}
corrPears_3.1 <- jmv::corrMatrix(
    data = df_PC3.1,
    pearson = TRUE,
    vars = vars(var1, var2),
    flag = TRUE,
    ci = TRUE,
    plots = TRUE,
    plotDens = TRUE,
    plotStats = TRUE)
corrPears_3.1$matrix
```

#### By changing the magnitude and adding some noise
```{r}
corrPears_3.2 <- jmv::corrMatrix(
    data = df_PC3.1,
    pearson = TRUE,
    vars = vars(var3, var4),
    flag = TRUE,
    ci = TRUE,
    plots = TRUE,
    plotDens = TRUE,
    plotStats = TRUE)
corrPears_3.2$matrix
```

## Test the correlation between three highly correlated continuous variables

```{r echo=FALSE, message=FALSE}

df_PC4 <- rawData %>%
  select(
    righthippocampus,
    lefthippocampus,
    leftententorhinalarea
    ) %>%
  drop_na() 
head(df_PC4)
```

```{r}
corrPears_4 <- jmv::corrMatrix(
    data = df_PC4,
    pearson = TRUE,
    vars = vars(righthippocampus, lefthippocampus, leftententorhinalarea),
    flag = TRUE,
    ci = TRUE,
    plots = TRUE,
    plotDens = TRUE,
    plotStats = TRUE)
corrPears_4$matrix
```

```{r}
corrPears_4$plot
```

## Test the correlation between two highly correlated continuous variables and one weakly correlated

```{r echo=FALSE, message=FALSE}

df_PC5 <- rawData %>%
  select(
    righthippocampus,
    lefthippocampus,
    opticchiasm
    ) %>%
  drop_na() 
head(df_PC5)
```

```{r}
corrPears_5 <- jmv::corrMatrix(
    data = df_PC5,
    pearson = TRUE,
    vars = vars(righthippocampus, lefthippocampus, opticchiasm),
    flag = TRUE,
    ci = TRUE,
    plots = TRUE,
    plotDens = TRUE,
    plotStats = TRUE)
corrPears_5$matrix
```

```{r}
corrPears_5$plot
```

## Test the correlation between two positively correlated continuous variables and one negatively correlated

```{r echo=FALSE, message=FALSE}

df_PC6 <- rawData %>%
  select(
    lefthippocampus,
    subjectage,
    opticchiasm
    ) %>%
  drop_na() 
head(df_PC6)
```

```{r}
corrPears_6 <- jmv::corrMatrix(
    data = df_PC6,
    pearson = TRUE,
    vars = vars(lefthippocampus, subjectage, opticchiasm),
    flag = TRUE,
    ci = TRUE,
    plots = TRUE,
    plotDens = TRUE,
    plotStats = TRUE)
corrPears_6$matrix
```

```{r}
corrPears_6$plot
```

#### Change predictor order

```{r}
corrPears_7 <- jmv::corrMatrix(
    data = df_PC6,
    pearson = TRUE,
    vars = vars(subjectage, lefthippocampus, opticchiasm),
    flag = TRUE,
    ci = TRUE,
    plots = TRUE,
    plotDens = TRUE,
    plotStats = TRUE)
corrPears_7$matrix
```







































