---
title: "T-Test testing"
output:
  html_document:
    df_print: paged
  pdf_document: default
---


This test series is intended to cover the three types of use of the T-Test without testing the limitations of these functions. However, we also test few different options that can influence the calculations. 
The complete list of parameters can be found here:
- INDEPENDENT SAMPLES T-TEST: https://www.jamovi.org/jmv/ttestis.html 
- PAIRED SAMPLES T-TEST: https://www.jamovi.org/jmv/ttestps.html
- ONE SAMPLE T-TEST: https://www.jamovi.org/jmv/ttestones.html
<br>
We propose to test only necessary options first knowing that some options not tested yet should be added and tested in the futur...

# Local calculation (single node)

## Setup environment and load dataset

```{r message=F}
library(tidyverse)
library(jmv)

df_TT <- read_csv("Data/desd-synthdata.csv", 
    col_types = cols(gender = col_factor(levels = c("M", 
        "F"))))
```


## Test #1: independent samples T-Test
In this test, we want to know if the volume of the female's left hippocampus is different from that of the male. 

- We check if the means are different.

```{r}
ttestIS(
  formula = lefthippocampus ~ gender, 
  data = df_TT,
  hypothesis = 'different',
  effectSize = T
  )
```


- We check if the mean volumes are greater for men than for women.

```{r}
ttestIS(
  formula = lefthippocampus ~ gender, 
  data = df_TT,
  hypothesis = 'oneGreater',
  meanDiff = T,
  ci = T,
  effectSize = T
  )
```



## Test #2: paired samples T-Test
In this test, we consider two paired continuous variables for which we want to know if their two means are different.

```{r}
ttestPS(
  data = df_TT, 
  pairs = list(list(i1 = 'lefthippocampus', i2 = 'righthippocampus')),
  effectSize = T
)
```

## Test #3: Comparison with a reference value
In this test, we compare the mean of the left and right hippocampus volumes separetely, with a reference value 3. 

```{r}
ttestOneS(
  data = df_TT, 
  vars = vars(lefthippocampus, righthippocampus),
  testValue = 3.)
```

# Federated calculation

In this section, we must repeat the previous tests by splitting the data set into three subsets of data of different sizes.



```{r}
dataset0 <- df_TT %>%
  select(c("lefthippocampus", "alzheimerbroadcategory", "gender", "agegroup")) %>%
  drop_na() 

dataset1 <- dataset0 %>%
  dplyr::slice(1:225) %>%
  mutate(dataset = "dataset1")
write_csv(dataset1, "Data/dataset1.csv")

dataset2 <- dataset0 %>%
  dplyr::slice(226:426) %>%
  mutate(dataset = "dataset2")
write_csv(dataset2, "Data/dataset2.csv")

dataset3 <- dataset0 %>%
  dplyr::slice(427:718) %>%
  mutate(dataset = "dataset3")
write_csv(dataset3, "Data/dataset3.csv")

dataset0 <- rbind(dataset1, dataset2, dataset3)
write_csv(dataset0, "Data/dataset_0.csv")

dataset0 %>%
  ggplot(aes(x = lefthippocampus, color = gender, fill = gender)) + 
  geom_histogram(bins = 30, alpha = 0.25, position = "identity") + 
  facet_grid(. ~ dataset) +
  theme_bw() 
```







